---
title: " "
author: "RawanAloula"
date: " "
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


<br>

#### Load data 

```{r}
library(nycflights13)
head(flights, 10)
dim(flights)
tail(flights, 10)
```

```{r}

colnames(flights)

```




<br>

#### 1. 

The function calculate basic statistics hence "Stat" and returns Min, Max, Median hence "3M"

<br>

```{r}

Stat.3M <- function(x){
  x <- x[!is.na(x)]
  a = min(x)
  b = max(x)
  c = median(x)
  sprintf( "min = %.2f   , max = %.2f   and median is %.2f", a, b, c)
}

```

```{r}
Stat.3M(flights$month)
```

<br>


#### 2. Write a function that categorizes time 

"Morning" for values from 5 am to 11:59 am 

"Afternoon" for values from 12 pm to 4:59 pm

"Evening" for values from 5 pm to 8:59 pm

"Night"  for values from 9 pm to 4:59 am


<br>

It is advised to use the name of the function as a verb hence "Categorize", and "t" to emphasize time categorization.

<br>

```{r}

Categorize.t <-  function(data, v){
  x <- data[[v]]
  x <- x[!is.na(x)]
  
  x[ (x >= 500 & x <= 1159 )] <- "Morning"
  x[ (x >= 1200 & x <= 1659 )] <- "Afternoon"
  x[ (x >= 1700 & x <= 2059 )] <- "Evening"
  x[ (x >= 2100 & x <= 4590 )] <- "Night"
  
  
  out <- cbind(data[[v]][!is.na(data[[v]])],x)
  return(out)
}
```


```{r}
head(Categorize.t(flights, "dep_time"),10)
tail(Categorize.t(flights, "dep_time"),10)
```

<br>

#### 3.  Find median of all columns in the flights dataset using a for loop.  


<br>


The main purpose of the function is to calculate the median for all numeric variables in the data hence, "Median" and we put n emphases that it is a calculation for the entire data frame with "Data".            
<br>

##### 3.1) using a for loop

```{r}

Median.Data <- function(data){

  med <- c()
  var.name <- c()
  
  for(i in names(data)) {
    
    x <- data[[i]]
    x <- x[!is.na(x)]
    
    if(is.numeric(x)) {
      
      med <- c(med, median(x))
      var.name <- c(var.name, i)
    }
  
  }

  return(cbind(var.name,med))
}


```



```{r}
Median.Data(flights)
```

<br>

##### 3.2) using "summarise_if"

```{r , message= FALSE}
library(tidyverse)
```



```{r, message= FALSE}

flights %>%
    summarise_if(is.numeric, median, na.rm = TRUE)

```









